---
title: "Where are the Same-Sex Households?"
output: html_document
---
Assignment 6 - Spatial Analysis     
Alex Cardelle     
October 18, 2021
```{r setup, include=FALSE, message=FALSE, echo=FALSE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
This map shows the spatial distribution of the U.S. population living in either a coupled opposite-sex or same-sex households. "Coupled" for this analysis means living in a household with either a married spouse or an unmarried partner present.

Unless otherwise noted, all information was collected from the 2019 American Community Survey (ACS 2019) through the Census API at <http://data.census.gov>.

**Code Used to Generate Map** 

```{r library, eval=TRUE, echo=TRUE, quiet=FALSE, message=FALSE, results='hide', include=FALSE}
library(leaflet)
library(sf)
library(tidyverse)
library(tidycensus)
library(ggthemes)
library(ggspatial)
library(htmlwidgets)
library(tidytransit)
```

``` {r ACS gay married map, eval=TRUE, echo=TRUE, quiet=TRUE, message=FALSE, results='hide', warning=FALSE}

#load the variables

# B09019_010
# Estimated total population in households with an opposite-sex spouse present
# B09019_012
# Estimate total population in households with an opposite-sex unmarried partner present
# B09019_011
# Estimated total population in households with a same-sex spouse present
# B09019_013
# Estimate total population in households with a same-sex unmarried partner present


gay_hh <- get_acs(geography = "county", year = 2019, survey = "acs5",
                    variables = c(tot_het_ma = "B09019_010",tot_het_um = "B09019_012",
                                  tot_gay_ma = "B09019_011", tot_gay_um = "B09019_013"),
                    output = "wide", geometry = TRUE) 

#mutate the data to eliminate of the margin of error

gay_hh <- gay_hh %>%
  select(-tot_het_maM, -tot_het_umM, -tot_gay_maM, -tot_gay_umM) %>%
  filter(tot_het_maE > 0) %>%
  mutate(pct_gay = round(100 * (tot_gay_maE + tot_gay_umE) / (tot_het_maE + tot_het_umE), 1))

#define the palette
household_palette <- colorNumeric(c("lightgrey", "darkblue"),
                                gay_hh$pct_gay)
#plot the map
gay_map <-
  leaflet(gay_hh) %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(fillColor = ~household_palette(pct_gay), weight = 1, color = "grey", fillOpacity = 0.8,
              highlightOptions = highlightOptions(fillColor = "yellow", fillOpacity = 0.9),
              label = gay_hh$NAME,
              popup = paste("Total population in coupled households: ", (gay_hh$tot_het_maE+gay_hh$tot_het_umE+gay_hh$tot_gay_maE+gay_hh$tot_gay_umE),
                            "<br/>",
                            "Total population in coupled same-sex households: ",
                            (gay_hh$tot_gay_maE+gay_hh$tot_gay_umE), " (", 
                            gay_hh$pct_gay, "%)",sep = "")) %>%
    addLegend("bottomright", pal = household_palette, values = ~pct_gay,
    title = "Percent of the U.S. population<br/> in households living in<br/>same-sex households",
    labFormat = labelFormat(suffix = "%"),
    opacity = 1)
```

**Interactive Map**
```{r plot the gays, eval=TRUE, echo=FALSE}
gay_map

```

``` {r limit the gays, echo=FALSE, eval=FALSE}

limits <- st_bbox(gay_hh)

gay_map2 <- gay_map %>%
   fitBounds( lng1 = as.numeric(limits[1]),
                lat1 = as.numeric(limits[2]),
                lng2 = as.numeric(limits[3]),
                lat2 = as.numeric(limits[4])) %>%
   setMaxBounds( lng1 = as.numeric(limits[1]),
                lat1 = as.numeric(limits[2]),
                lng2 = as.numeric(limits[3]),
                lat2 = as.numeric(limits[4])) 

gay_map2
```

Produced at the Harvard GSD.